---
title: "Tooth Growth Analysis"
author: <a href="http://www.ryantillis.com"> Ryan Tillis </a>
date: "August 9, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

This is a brief graphical and numerical exploration of a data set. The goal is to summarize the data and perform some confidence interval/hypothesis tests.

## Loading the data

```{r beginning}
library(datasets)
library(ggplot2)
library(RColorBrewer)
library(grDevices)

data(ToothGrowth)
attach(ToothGrowth)

#A first look shows us that there are 3 variables and 60 observations
str(ToothGrowth)

#Shows that the dose is in .5 increments with 3 unique levels - converted to factors
summary(ToothGrowth)
unique(ToothGrowth$dose)
ToothGrowth$dose<-as.factor(ToothGrowth$dose)

```

## Graphical Analysis

In this section we explore the relationship between dose size, dose type and tooth length.

```{r pressure, fig.height=3, echo=FALSE}
require(gridExtra)

theme <- theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) 

plot1 <- ggplot(aes(x = factor(dose), y = len), data = ToothGrowth) + geom_boxplot(aes(fill = factor(dose))) + theme + scale_fill_brewer(palette="PuBu") + labs(title = "Figure 1")

plot2 <- ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + theme + scale_fill_brewer(palette="YlOrRd") + labs(title = "Figure 2")

grid.arrange(plot1, plot2, ncol=2)
```

**Figure 1** shows that as doseage increases so does tooth length.

```{r plot2, fig.height = 3, echo = FALSE}
ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + facet_wrap(~ dose) + theme + scale_fill_brewer(palette="YlOrRd")+ labs(title = "Figure 3") + annotate("text", x = 1.5, y = 2, label = c("p = .03","p = .0005","p = .5"))
```

**Figure 2** suggests that Orange Juice is generally more effective than vitamin C across all dose levels.  

In **Figure 3** things get a little more interesting. For the first 2 doseage ranges, Orange Juice appears marginally more effective than vitamin C at stimulating tooth growth. For the 3rd dose it appears a push.

**Figure 3** shows that the difference between the first 2 dose ranges appears significant but at the highest dose of .5, the two supplements are equally effective. P-values attained from the numerical analysis section below are annotated onto the graph.

## Conclusions

####Reformatting Data

The first task is to transform the data into a useful format for computing confidence intervals. To do this I used the split funciton. From there the relevant columns can be pulled out of the list to perform t.tests.

```{r split}
#This operation splits the data frame by dose and supplement type. The result is a list 6 data frames each with 10 observations controlled for dose and supplement type.

split_tooth <- split(ToothGrowth, f = list(ToothGrowth$dose, ToothGrowth$supp))
```

Aggregation shows the means and sd to be compared
```{r agg1}
aggregate(len,list(supp,dose),mean)
aggregate(len,list(supp,dose),sd)
```

####Computing the Confidence Intervals
The following conclusions can be stated with 95% confidence levels. P-values are displayed below. Codes for the t.test are included in the appendix.

#####Is Orange Juice more effective than vitamin C across doses? <span style="color:green">YES</span>
```{r t1, echo = FALSE}
t.test(c(split_tooth[[1]]$len,split_tooth[[2]]$len,split_tooth[[3]]$len),c(split_tooth[[4]]$len,split_tooth[[5]]$len,split_tooth[[6]]$len), alternative = "greater")$p.value
```

#####Is Orange Juice more effective than Vitamin C for a .5 dose? <span style="color:green">YES</span>
```{r t2, echo = FALSE}
t.test(split_tooth[[1]]$len,split_tooth[[4]]$len, alternative = "greater")$p.value
```
#####Is Orange Juice more effective than Vitamin C for a 1 dose? <span style="color:green">YES</span>
```{r t3, echo = FALSE}
t.test(split_tooth[[2]]$len,split_tooth[[5]]$len, alternative = "greater")$p.value
```
#####Is Orange Juice more effective than Vitamin C for a 2 dose? <span style="color:red">INCONCLUSIVE</span>
```{r t4, echo = FALSE}
t.test(split_tooth[[3]]$len,split_tooth[[6]]$len, alternative = "greater")$p.value
```
#####Is a 2 dose (any supp) more effective than a .5 dose? <span style="color:green">YES</span>
```{r t5, echo = FALSE}
t.test(c(split_tooth[[1]]$len,split_tooth[[4]]$len),c(split_tooth[[2]]$len,split_tooth[[5]]$len), alternative = "less")$p.value
```
#####Is a 2 dose more (any supp) effective than a .5 dose? <span style="color:green">YES</span>
```{r t6, echo = FALSE}
t.test(c(split_tooth[[1]]$len,split_tooth[[4]]$len),c(split_tooth[[3]]$len,split_tooth[[6]]$len), alternative = "less")$p.value
```
#####Is a 2 dose more (any supp) effective than a .5 dose? <span style="color:green">YES</span>
```{r t7, echo = FALSE}
t.test(c(split_tooth[[2]]$len,split_tooth[[5]]$len),c(split_tooth[[3]]$len,split_tooth[[6]]$len), alternative = "less")$p.value
```

###Appendix - Plot and Conf Int Codes

```{r app, eval = FALSE}
#Is Orange Juice more effective than vitamin C across doses? <span style="color:green">YES</span>
t.test(c(split_tooth[[1]]$len,split_tooth[[2]]$len,split_tooth[[3]]$len),c(split_tooth[[4]]$len,split_tooth[[5]]$len,split_tooth[[6]]$len), alternative = "greater")$p.value

#Is Orange Juice more effective than Vitamin C for .5 dose? YES
t.test(split_tooth[[1]]$len,split_tooth[[4]]$len, alternative = "greater")$p.value

#Is Orange Juice more effective than Vitamin C for 1 dose? YES
t.test(split_tooth[[2]]$len,split_tooth[[5]]$len, alternative = "greater")$p.value

#Is Orange Juice more effective than Vitamin C for 2 dose? Inconclusive
t.test(split_tooth[[3]]$len,split_tooth[[6]]$len, alternative = "greater")$p.value

#Is 2 dose more effective that .5 dose? YES
t.test(c(split_tooth[[1]]$len,split_tooth[[4]]$len),c(split_tooth[[2]]$len,split_tooth[[5]]$len), alternative = "less")$p.value

#Is 2 dose more effective that .5 dose? YES
t.test(c(split_tooth[[1]]$len,split_tooth[[4]]$len),c(split_tooth[[3]]$len,split_tooth[[6]]$len), alternative = "less")$p.value

#Is 2 dose more effective that .5 dose? YES
t.test(c(split_tooth[[2]]$len,split_tooth[[5]]$len),c(split_tooth[[3]]$len,split_tooth[[6]]$len), alternative = "less")$p.value
```

```{r pressureA, fig.height=3, eval=FALSE}
require(gridExtra)

theme <- theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) 

#FIGURE 1
plot1 <- ggplot(aes(x = factor(dose), y = len), data = ToothGrowth) + geom_boxplot(aes(fill = factor(dose))) + theme + scale_fill_brewer(palette="PuBu") + labs(title = "Figure 1")

#FIGURE 2
plot2 <- ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + theme + scale_fill_brewer(palette="YlOrRd") + labs(title = "Figure 2")

grid.arrange(plot1, plot2, ncol=2)
```

```{r plot2A, fig.height = 3, eval = FALSE}
#FIGURE 3
ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + facet_wrap(~ dose) + theme + scale_fill_brewer(palette="YlOrRd")+ labs(title = "Figure 3") + annotate("text", x = 1.5, y = 2, label = c("p = .03","p = .0005","p = .5"))
```
Check out my website at: <http://www.ryantillis.com/>
